# ANMI018 spatial analysis

```{r load_libraries, comment="", message=FALSE, warning=FALSE}
library(tidyverse)
library(vegan)
library(adespatial)
library(ade4)
library(zCompositions)
library(ape)
library(spdep)
library(lmPerm)
library(tinytable)
select <- dplyr::select

```

```{r load_data, comment="", message=FALSE, warning=FALSE}
comm_data <- read_tsv("data/genome_counts_filt.tsv")
metadata <- read_tsv("data/final_combined_stats.tsv") %>%
  filter(cryosection %in% c("G121eO301A","G121eO302A","G103bI301A","G103bI301B","G121eI103A","G121eI103B","G121eI104C"))
```

```{r prepare_data, comment="", message=FALSE, warning=FALSE}
genome<-comm_data$genome
comm_data<-t(comm_data[,-1])
colnames(comm_data)<-genome
comm_data<-comm_data[rowSums(comm_data)>0,]

# comm data and metadata in the same order

metadata<-metadata[metadata$microsample%in%rownames(comm_data),]
comm_data<-data.frame(comm_data[rownames(comm_data)%in%metadata$microsample,])
```

### Explore study design

```{r explore_design, comment="", message=FALSE, warning=FALSE}
# Colon and Caecum samples
table(metadata$section,metadata$cryosection) %>% 
  as.data.frame() %>% 
  pivot_wider(names_from="Var2",values_from = "Freq") %>% 
  tt()

comm_ca<-comm_data[metadata$section=="Caecum right",]
comm_ca<-comm_ca[,colSums(comm_ca)>0]
comm_co<-comm_data[metadata$section=="Colon",]
comm_co<-comm_co[,colSums(comm_co)>0]
comm_data<-comm_data[,colSums(comm_data)>0]

metadata_ca<-metadata[metadata$section=="Caecum right",]
metadata_co<-metadata[metadata$section=="Colon",]

table(metadata$section,metadata$cryosection) %>% 
  as.data.frame() %>% 
  pivot_wider(names_from="Var2",values_from = "Freq") %>% 
  tt()
```


### Transform the data for posterior multivariate analyses

```{r transform, comment="", message=FALSE, warning=FALSE}
comm_ca_zeroRepl <- cmultRepl(comm_ca, method = "GBM", output = "prop",z.warning = 0.95)
comm_co_zeroRepl <- cmultRepl(comm_co, method = "GBM", output = "prop",z.warning = 0.95)

metadata_ca_clr<-metadata_ca[metadata_ca$microsample%in%rownames(comm_ca_zeroRepl),]
metadata_co_clr<-metadata_co[metadata_co$microsample%in%rownames(comm_co_zeroRepl),]

clr_transform <- function(x) {log(x) - mean(log(x), na.rm = TRUE)}
comm_ca_clr <- t(apply(comm_ca_zeroRepl, 1, clr_transform)) 
comm_co_clr <- t(apply(comm_co_zeroRepl, 1, clr_transform)) 
```

## Diversity partitioning

```{r diversity_partitioning, comment="", message=FALSE, warning=FALSE}
ca_adipart<-adipart(as.matrix(comm_ca>0)*1,as.matrix(metadata_ca[,c("microsample","cryosection","animal")]),
                    index="richness",weights="unif",relative=FALSE)

co_adipart<-adipart(as.matrix(comm_co>0)*1,as.matrix(metadata_co[,c("microsample","cryosection","animal")]),
                    index="richness",weights="unif",relative=FALSE)

ca<-cbind(c(ca_adipart$oecosimu$mean[1],ca_adipart$statistic[1]),
          c(ca_adipart$oecosimu$mean[4],ca_adipart$statistic[4]),
          c(ca_adipart$oecosimu$mean[2],ca_adipart$statistic[2]),
          c(ca_adipart$oecosimu$mean[5],ca_adipart$statistic[5]),
          c(ca_adipart$oecosimu$mean[3],ca_adipart$statistic[3])
)
rownames(ca)<-c("observed","expected")
colnames(ca)<-c("alpha.1","beta.1","alpha.2","beta.2","gamma")

co<-cbind(c(co_adipart$oecosimu$mean[1],co_adipart$statistic[1]),
          c(co_adipart$oecosimu$mean[4],co_adipart$statistic[4]),
          c(co_adipart$oecosimu$mean[2],co_adipart$statistic[2]),
          c(co_adipart$oecosimu$mean[5],co_adipart$statistic[5]),
          c(co_adipart$oecosimu$mean[3],co_adipart$statistic[3])
)
rownames(co)<-c("observed","expected")
colnames(co)<-c("alpha.1","beta.1","alpha.2","beta.2","gamma")


par(mfrow=c(1,2))
barplot(ca, beside = TRUE,
        col = c("black", "white"))
title(main = "Adipart caecum", font.main = 4)
barplot(co, beside = TRUE,
        col = c("black", "white"),
        legend = rownames(co))
title(main = "Adipart colon", font.main = 4)
par(mfrow=c(1,1))
```

### Sample distributions between animals and cryosections

```{r sample_distribution, comment="", message=FALSE, warning=FALSE}
table(metadata_co$animal,metadata_co$cryosection) %>% 
  as.data.frame() %>% 
  pivot_wider(names_from="Var2",values_from = "Freq") %>% 
  tt()
table(metadata_ca$animal,metadata_ca$cryosection) %>% 
  as.data.frame() %>% 
  pivot_wider(names_from="Var2",values_from = "Freq") %>% 
  tt()

plot(metadata_co$Xcoord,metadata_co$Ycoord)

plot(metadata_ca$Xcoord,metadata_ca$Ycoord)
```

### Explore community data

#### Caecum

```{r caecum_nmds, comment="", message=FALSE, warning=FALSE}
nmds_ca<-comm_ca_clr%>%
  as.data.frame()%>%
  rownames_to_column(var="microsample")%>%
  bind_cols(metadata_ca_clr)%>%
  select(contains("bin"))%>%
  metaMDS(.,trymax = 400,distance="euclidean",autotransform = FALSE, trace=0)%>%
  vegan::scores(.,display = "sites") %>%
  as_tibble(., rownames = "microsample") %>%
  bind_cols(metadata_ca_clr)%>%
  group_by(cryosection) %>%
  mutate(microsample_x=mean(NMDS1), microsample_y=mean(NMDS2))

nmds_ca %>%
  ggplot(aes(x=NMDS1,y=NMDS2, color=cryosection,shape=animal)) +
  geom_point(size=2) +
  geom_segment(aes(x=microsample_x, y=microsample_y, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  scale_color_manual(values=c("red","blue","black","gray","green")) +
  theme_classic() +
  theme(legend.position="right", legend.box="vertical") +
  guides(color=guide_legend(title="Cryosection"))
```

#### Colon 

```{r colon_nmds, comment="", message=FALSE, warning=FALSE}
nmds_co<-comm_co_clr%>%
  as.data.frame()%>%
  rownames_to_column(var="microsample")%>%
  bind_cols(metadata_co_clr)%>%
  select(contains("bin"))%>%
  metaMDS(.,trymax = 400,k=3,distance="euclidean",autotransform = FALSE, trace=0)%>%
  vegan::scores(.,display = "sites") %>%
  as_tibble(., rownames = "microsample") %>%
  bind_cols(metadata_co_clr)%>%
  group_by(cryosection) %>%
  mutate(microsample_x=mean(NMDS1), microsample_y=mean(NMDS2))

nmds_co %>%
  ggplot(aes(x=NMDS1,y=NMDS2, color=cryosection,shape=animal)) +
  geom_point(size=2) +
  geom_segment(aes(x=microsample_x, y=microsample_y, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  scale_color_manual(values=c("magenta","orange")) +
  theme_classic() +
  theme(legend.position="right", legend.box="vertical") +
  guides(color=guide_legend(title="Cryosection"))
```

## Spatial analysis

### Caecum

```{r caecum_spatial_prepare, comment="", message=FALSE, warning=FALSE}
comm_ca_clr_G121eI104C<-comm_ca_clr%>%
  as.data.frame()%>%
  rownames_to_column(var="microsample")%>%
  bind_cols(metadata_ca_clr)%>%
  filter(cryosection=="G121eI104C")%>%
  filter(!is.na(Xcoord))%>%
  select(contains("bin"))
metadata_ca_G121eI104C<-metadata_ca_clr%>%
  filter(!is.na(Xcoord))%>%
  filter(cryosection=="G121eI104C")
```

#### Mantel test

```{r caecum_spatial_mantel, comment="", message=FALSE, warning=FALSE}
mantel(dist(metadata_ca_G121eI104C[,c("Xcoord","Ycoord")]),dist(comm_ca_clr_G121eI104C))
```

#### Distance decay plot

```{r caecum_spatial_decay, comment="", message=FALSE, warning=FALSE}
toplot<-data.frame(spat_dist=as.numeric(dist(metadata_ca_G121eI104C[,c("Xcoord","Ycoord")])),
                   comm_dist=as.numeric(dist(comm_ca_clr_G121eI104C)))
ggplot(toplot,aes(x=spat_dist,y=comm_dist))+
  geom_point()+
  geom_smooth()
ggplot(toplot,aes(x=spat_dist,y=comm_dist))+
  geom_smooth()
anova(lmp(comm_dist~spat_dist,data=toplot))
```

#### Complex spatial structures
```{r caecum_spatial_complex, comment="", message=FALSE, warning=FALSE}
set.seed(111)
G121eI104C_spat_SWNs<-listw.candidates(data.frame(metadata_ca_G121eI104C[,c("Xcoord","Ycoord")]),
                                       nb = c("rel"), 
                                       weights = c("fdown", "fup"), 
                                       y_fdown = c(2,5), y_fup = c(0.1,0.5))
G121eI104C_W_sel <- listw.select(comm_ca_clr_G121eI104C, G121eI104C_spat_SWNs, MEM.autocor = "positive", 
                                 p.adjust = T, method="FWD")
G121eI104C_W_sel$candidates
```


### Colon

```{r colon_spatial_prepare, comment="", message=FALSE, warning=FALSE}
comm_co_clr_G121eO301A<-comm_co_clr%>%
  as.data.frame()%>%
  rownames_to_column(var="microsample")%>%
  bind_cols(metadata_co_clr)%>%
  filter(cryosection=="G121eO301A")%>%
  select(contains("bin"))
metadata_co_G121eO301A<-metadata_co_clr%>%
  filter(cryosection=="G121eO301A")
```

#### Mantel test

```{r colon_spatial_mantel, comment="", message=FALSE, warning=FALSE}
mantel(dist(metadata_co_G121eO301A[,c("Xcoord","Ycoord")]),dist(comm_co_clr_G121eO301A))
```

#### Distance decay plot
```{r colon_spatial_decay, comment="", message=FALSE, warning=FALSE}

toplot<-data.frame(spat_dist=as.numeric(dist(metadata_co_G121eO301A[,c("Xcoord","Ycoord")])),
                   comm_dist=as.numeric(dist(comm_co_clr_G121eO301A)))
ggplot(toplot,aes(x=spat_dist,y=comm_dist))+
  geom_point()+
  geom_smooth()
ggplot(toplot,aes(x=spat_dist,y=comm_dist))+
  geom_smooth()
anova(lmp(comm_dist~spat_dist,data=toplot))
```

#### Complex spatial structures

```{r colon_spatial_complex, comment="", message=FALSE, warning=FALSE}
set.seed(111)
G121eO301A_spat_SWNs<-listw.candidates(data.frame(metadata_co_G121eO301A[,c("Xcoord","Ycoord")]),
                                       nb = c("rel"), 
                                       weights = c("fdown", "fup"), 
                                       y_fdown = c(2,5), y_fup = c(0.1,0.5))
G121eO301A_W_sel <- listw.select(comm_co_clr_G121eO301A, G121eO301A_spat_SWNs, MEM.autocor = "positive", 
                                 p.adjust = T, method="FWD")
G121eO301A_W_sel$candidates
G121eO301A_W_sel$best$summary

par(mfrow=c(2,2))
for(i in 1:ncol(G121eO301A_W_sel$best$MEM.select)){
  s.value(metadata_co_G121eO301A[,c("Xcoord","Ycoord")],
          G121eO301A_W_sel$best$MEM.select[,i], 
          sub=G121eO301A_W_sel$best$summary$variables[i], 
          csub=2,
          include.origin = F,
          csize=0.8)
}
par(mfrow=c(1,1))
```

```{r colon_spatial_complex_plot, comment="", message=FALSE, warning=FALSE}
s.value(metadata_co_G121eO301A[,c("Xcoord","Ycoord")],
        rda(comm_co_clr_G121eO301A,
            G121eO301A_W_sel$best$MEM.select)$CCA$u[,1],
        sub=c("Spatial patterns of Colon"), 
        csub=2,
        include.origin = F,
        csize=0.8)

RsquareAdj(rda(comm_co_clr_G121eO301A,
               G121eO301A_W_sel$best$MEM.select))

# # Separating coarse- and fine-grain structures
# g.MEM.broad<-G121eO301A_W_sel$best$MEM.select[,c(4)]
# g.MEM.fine<-G121eO301A_W_sel$best$MEM.select[,c(1:3)]
# 
# par(mfrow=c(1,2))
# # Coarse-grain MEMs map
# s.value(metadata_co_G121eO301A[,c("Xcoord","Ycoord")],
#         rda(comm_co_clr_G121eO301A,
#             g.MEM.broad,cbind(g.MEM.fine))$CCA$u[,1],
#         sub=c("Coarse-grain"), 
#         csub=2,
#         include.origin = F,
#         csize=0.8)
# # Fine-grain MEMs map
# s.value(metadata_co_G121eO301A[,c("Xcoord","Ycoord")],
#         rda(comm_co_clr_G121eO301A,
#             g.MEM.fine,cbind(g.MEM.broad))$CCA$u[,1],
#         sub=c("Fine-grain"), 
#         csub=2,
#         include.origin = F,
#         csize=0.8)
# 
# RsquareAdj(rda(comm_co_clr_G121eO301A,
#                g.MEM.broad,
#                cbind(g.MEM.fine)))
# RsquareAdj(rda(comm_co_clr_G121eO301A,
#                g.MEM.fine,
#                cbind(g.MEM.broad)))
```

## rlqESLTP of the Colon cryosection

```{r load_data_rlqESLTP, comment="", message=FALSE, warning=FALSE}
load("data/data.RData")
source("data/JEC_1743_sm_apps5.txt")
```

```{r prepare_data_rlqESLTP, comment="", message=FALSE, warning=FALSE}
comm_co_G121eO301A<-comm_co%>%
  as.data.frame()%>%
  rownames_to_column(var="microsample")%>%
  bind_cols(metadata_co)%>%
  filter(cryosection=="G121eO301A")%>%
  select(contains("bin"))
metadata_co_G121eO301A<-metadata_co%>%
  filter(cryosection=="G121eO301A")

comp <- decostand(comm_co_G121eO301A,MARGIN=1,method="hellinger")	# Standardize composition by rows
colnames(comp)<-gsub("\\.",":",names(comp))
env <- data.frame(log_seq_counts=log(metadata_co_G121eO301A$fastp_total_after_filtering.total_bases),
                  div=rowSums(comp>0))			   			# Environmental matrix

genome_gifts<-genome_gifts[rownames(genome_gifts)%in%colnames(comp),]
genome_funct <- genome_gifts %>%
  as.data.frame() %>%
  rownames_to_column(var = "id") %>%                         # Preserve row names as a column
  pivot_longer(-id, names_to = "column", values_to = "value") %>%  # Reshape to long format
  mutate(group = substr(column, 1, 3)) %>%                  # Extract the first 3 characters of column names
  group_by(id, group) %>%                                   # Group by row (id) and prefix
  summarise(mean_value = mean(value, na.rm = TRUE), .groups = "drop") %>% # Calculate rowMeans for each group
  pivot_wider(names_from = group, values_from = mean_value) %>%  # Reshape back to wide format
  column_to_rownames(var = "id") 
genome_funct<-genome_funct[,-c(19:21)]

phy <- genome_tree				       		# Phylogenetic tree constructed with Phylomatic V3 + Phylocom

phy <- drop.tip(phy, setdiff(phy$tip.label, rownames(genome_funct)))
spa <- metadata_co_G121eO301A[,c("Xcoord","Ycoord")]				       		# Matrix of X Y spatial coordinates	
comp<-comp[,match(phy$tip.label,colnames(comp))]
genome_funct<-genome_funct[match(phy$tip.label,rownames(genome_funct)),]

mean(phy$tip.label==colnames(comp))

# Change the phylogenetic tree (an object of the class "phylo") into an object of the class "phylog", used by ade4
phylog <- newick2phylog(write.tree(phy)) # => "phylog" class
colnames(comp)<-gsub(":","_",colnames(comp))
rownames(genome_funct)<-gsub(":","_",rownames(genome_funct))
```

### Correspondence analysis of the composition matrix 

```{r correspondence_rlqESLTP, comment="", message=FALSE, warning=FALSE}
coacomp <- dudi.coa(comp, scan = FALSE, nf = 108)
summary(coacomp)
```

### Spatial analysis 

#### Gabriel graph

```{r gabriel_graph, comment="", message=FALSE, warning=FALSE}
nb1 <- graph2nb(gabrielneigh(as.matrix(spa)), sym = T)
nb1

lw1 <- nb2listw(nb1) # gives a neighbours list with spatial weights (i.e. matrix W)
lw1
```

```{r gabriel_graph_plot, comment="", message=FALSE, warning=FALSE}
par(mfrow=c(1,1))
plot(nb1, spa, pch = 21, bg = 'red')
title(main="Gabriel Graph")
class(nb1)
# [1] "nb"
# This is the Gabriel graph -the graph that in this case defines which points 
# are connected

# The matrix of spatial variables is obtained as the eigenvectors of a
# neighbour matrix. This matrix is analysed by principal component analysis.
nb1.neigh <- nb2neig(nb1)
vecspa <- scores.neig(nb1.neigh)
pcaspa <- dudi.pca(vecspa, coacomp$lw, scan = FALSE, nf = 199)
summary(pcaspa)
```

#### PCA analysis of the env matrix
```{r env_pca, comment="", message=FALSE, warning=FALSE}
pcaenv <- dudi.pca(env, row.w=coacomp$lw, scannf = FALSE, nf = 2)
summary(pcaenv)
```

#### The distances between species based on their biological traits, analyzed by PCoA.

```{r distances_traits, comment="", message=FALSE, warning=FALSE}
# Distance matrices for traits separately
listdis <- ldist.ktab(ktab.list.df(list(genome_funct)), c("Q"), scan = FALSE)
# choose 1
names(listdis)
summary(listdis)

# Distance matrix for traits together
disT <- dist.ktab(ktab.list.df(list(genome_funct)), c("Q"), scan = FALSE)
# choose 1
pcotraits <- dudi.pco(disT, coacomp$cw, full = TRUE)
summary(pcotraits)
```

#### The distances between species based on their phylogenetic relatedness, analysed by PCoA.

```{r distances_phylogeny, comment="", message=FALSE, warning=FALSE}
pcophy <- dudi.pco(as.dist(as.matrix(phylog$Wdist)[names(comp),names(comp)]),coacomp$cw, full = TRUE)
summary(pcophy)

## 11. Tests for phylogenetic signals in traits
## *******************************************
#Function rtest.decdiv() in appendix S5, Pavoine et al., 2011. [1]
#All traits together
phystot <- rtest.decdiv(phylog, rep(1, 108), as.dist(as.matrix(disT)[names(phylog$leaves),names(phylog$leaves)]), 
                        nrep = 99, vranking = "droot", optiontest = "less",ties.method = "average", option = 3)
phystot
```

#### Extended RLQ analysis

```{r extended_rlq, comment="", message=FALSE, warning=FALSE}
rlqmix <- rlqESLTP(pcaenv, pcaspa, coacomp, pcotraits, pcophy, scan = F, nf = 2)
barplot(rlqmix$eig)
rlqmix$eig[1]/sum(rlqmix$eig)
# [1] 0.74 
rlqmix$eig[2]/sum(rlqmix$eig)
# [1] 0.11   
```

**1st axis**

```{r extended_rlq_axis1, comment="", message=FALSE, warning=FALSE}
plot(rlqmix, xy=spa, ax=1, wh="S")
plot(rlqmix, phy=phylog, ax=1, wh="P")
plot(rlqmix, traits=genome_funct, ax=1, type="Q", wh="T")
plot(rlqmix, env=pcaenv$tab , ax=1, type="Q", wh="E")
```

**2nd axis**

```{r extended_rlq_axis2, comment="", message=FALSE, warning=FALSE}
plot(rlqmix, xy=spa, ax=2, wh="S")
plot(rlqmix, phy=phylog, ax=2, wh="P")
plot(rlqmix, traits=genome_funct, ax=2, type="Q", wh="T")
plot(rlqmix, env=pcaenv$tab , ax=2, type="Q", wh="E")
```