# ANMI018 spatial analysis

```{r load_libraries, comment="", message=FALSE, warning=FALSE}
library(tidyverse)
library(vegan)
library(adespatial)
library(ade4)
library(zCompositions)
library(ape)
library(spdep)
library(lmPerm)
library(tinytable)
select <- dplyr::select

```

```{r load_data, comment="", message=FALSE, warning=FALSE}
comm_data <- read_tsv("data/genome_counts_filt.tsv")
metadata <- read_tsv("data/alpha_div_filtered.tsv") %>% 
  filter(
    batch %in% c("MSEB0006", "MSEB0009", "MSEB0010", "MSEB0011", "MSEB0012", "MSEB0014", "MSEB0015"),
    type_binomial %in% c("positive"),
    genome_count_filtering30 %in% c("retained"),
    !section %in% "Ileum",
    !is.na(Xcoord))

metadata<-metadata %>%
  filter(!animal %in% "G103b")
```

```{r prepare_data, comment="", message=FALSE, warning=FALSE}
genome<-comm_data$genome
comm_data<-t(comm_data[,-1])
colnames(comm_data)<-genome
comm_data<-comm_data[rowSums(comm_data)>0,]

# comm data and metadata in the same order

metadata<-metadata[metadata$microsample%in%rownames(comm_data),]
comm_data<-data.frame(comm_data[rownames(comm_data)%in%metadata$microsample,])

mean(rownames(comm_data)==metadata$microsample)
```

### Explore study design

```{r explore_design, comment="", message=FALSE, warning=FALSE}
table(metadata$section,metadata$cryosection)

comm_ca<-comm_data[metadata$section=="Caecum right",]
comm_ca<-comm_ca[,colSums(comm_ca)>0]
comm_co<-comm_data[metadata$section=="Colon",]
comm_co<-comm_co[,colSums(comm_co)>0]
comm_data<-comm_data[,colSums(comm_data)>0]

metadata_ca<-metadata[metadata$section=="Caecum right",]
metadata_co<-metadata[metadata$section=="Colon",]

table(metadata$section,metadata$cryosection)
table(metadata$section,metadata$animal)
```


### Transform the data for posterior multivariate analyses

```{r transform, comment="", message=FALSE, warning=FALSE}
comm_ca_zeroRepl <- cmultRepl(comm_ca, method = "GBM", output = "prop",z.warning = 0.95)
comm_co_zeroRepl <- cmultRepl(comm_co, method = "GBM", output = "prop",z.warning = 0.95)

metadata_ca_clr<-metadata_ca[metadata_ca$microsample%in%rownames(comm_ca_zeroRepl),]
metadata_co_clr<-metadata_co[metadata_co$microsample%in%rownames(comm_co_zeroRepl),]

clr_transform <- function(x) {log(x) - mean(log(x), na.rm = TRUE)}
comm_ca_clr <- data.frame(t(apply(comm_ca_zeroRepl, 1, clr_transform))) 
comm_co_clr <- data.frame(t(apply(comm_co_zeroRepl, 1, clr_transform))) 
```

## Diversity partitioning

```{r diversity_partitioning, comment="", message=FALSE, warning=FALSE}
## Diversity partitioning in caecum and colon

ca_adipart<-adipart(as.matrix(comm_ca>0)*1,as.matrix(metadata_ca[,c("microsample","cryosection")]),
                    index="richness",weights="unif",relative=FALSE,nsimul=999,method="r2dtable")

co_adipart<-adipart(as.matrix(comm_co>0)*1,as.matrix(metadata_co[,c("microsample","cryosection")]),
                    index="richness",weights="unif",relative=FALSE,nsimul=999,method="r2dtable")

ca<-cbind(c(ca_adipart$oecosimu$mean[1],ca_adipart$statistic[1]),
          c(ca_adipart$oecosimu$mean[3],ca_adipart$statistic[3]),
          c(ca_adipart$oecosimu$mean[2],ca_adipart$statistic[2])
)
rownames(ca)<-c("observed","expected")
colnames(ca)<-c("alpha","beta","gamma")

co<-cbind(c(co_adipart$oecosimu$mean[1],co_adipart$statistic[1]),
          c(co_adipart$oecosimu$mean[3],co_adipart$statistic[3]),
          c(co_adipart$oecosimu$mean[2],co_adipart$statistic[2])
)
rownames(co)<-c("observed","expected")
colnames(co)<-c("alpha","beta","gamma")


par(mfrow=c(1,2))
bp<-barplot(ca, beside = TRUE,
        col = c("black", "white"),
        ylim=c(0,110))
text(x = apply(bp,2,mean), y = apply(ca,2,max)+7, labels = "***")
title(main = "Caecum", font.main = 4)
bp<-barplot(co, beside = TRUE,
        col = c("black", "white"),
        ylim=c(0,110))
text(x = apply(bp,2,mean), y = apply(co,2,max)+7, labels = "***")
title(main = "Colon", font.main = 4)
par(mfrow=c(1,1))
```

### Sample distributions between animals and cryosections

```{r sample_distribution, comment="", message=FALSE, warning=FALSE}
table(metadata_co$animal,metadata_co$cryosection)
table(metadata_ca$animal,metadata_ca$cryosection)

plot(metadata_co$Xcoord,metadata_co$Ycoord)

plot(metadata_ca$Xcoord,metadata_ca$Ycoord)
```
## Spatial analysis

### Caecum

```{r caecum_spatial_prepare, comment="", message=FALSE, warning=FALSE}
comm_ca_clr_G121eI104C <- comm_ca_clr %>%
  as.data.frame() %>%
  rownames_to_column(var="microsample") %>%
  left_join(metadata_ca_clr,by="microsample") %>%
  filter(cryosection=="G121eI104C") %>%
  filter(!is.na(Xcoord)) %>%
  column_to_rownames(var="microsample") %>% 
  select(contains("bin_"))

metadata_ca_G121eI104C <- metadata_ca_clr %>%
  filter(!is.na(Xcoord)) %>%
  filter(cryosection=="G121eI104C")
```

#### Mantel correlogram

```{r caecum_spatial_mantel, comment="", message=FALSE, warning=FALSE}
correlog_ca<-mantel.correlog(D.eco=dist(comm_ca_clr_G121eI104C),D.geo=dist(metadata_ca_G121eI104C[,c("Xcoord","Ycoord")]),nperm=999)
plot(correlog_ca)
```

#### Distance decay plot

```{r caecum_spatial_decay, comment="", message=FALSE, warning=FALSE}
toplot<-data.frame(spat_dist=as.numeric(dist(metadata_ca_G121eI104C[,c("Xcoord","Ycoord")])),
                   comm_dist=as.numeric(dist(comm_ca_clr_G121eI104C)))
ggplot(toplot,aes(x=spat_dist,y=comm_dist))+
  geom_point()+
  geom_smooth()
ggplot(toplot,aes(x=spat_dist,y=comm_dist))+
  geom_smooth()
summary(lmp(comm_dist~spat_dist,data=toplot))
```

#### Complex spatial structures
```{r caecum_spatial_complex, comment="", message=FALSE, warning=FALSE}
set.seed(111)
G121eI104C_spat_SWNs<-listw.candidates(data.frame(metadata_ca_G121eI104C[,c("Xcoord","Ycoord")]),
                                       nb = c("gab"), 
                                       weights = c("fdown", "fup"), 
                                       y_fdown = c(2,5), y_fup = c(0.1,0.5))
G121eI104C_W_sel <- listw.select(comm_ca_clr_G121eI104C, G121eI104C_spat_SWNs, MEM.autocor = "positive", 
                                 p.adjust = T, method="FWD")
G121eI104C_W_sel$candidates
```


### Colon

```{r colon_spatial_prepare, comment="", message=FALSE, warning=FALSE}
comm_co_clr_G121eO301A <- comm_co_clr %>%
  as.data.frame() %>%
  rownames_to_column(var="microsample") %>%
  left_join(metadata_co_clr,by="microsample") %>%
  filter(cryosection=="G121eO301A") %>%
  filter(!is.na(Xcoord)) %>%
  column_to_rownames(var="microsample") %>% 
  select(contains("bin_"))


metadata_co_G121eO301A <- metadata_co_clr %>%
  filter(!is.na(Xcoord)) %>%
  filter(cryosection=="G121eO301A")

plot(metadata_co_G121eO301A[,c("Xcoord","Ycoord")])
```

#### Mantel correlogram

```{r colon_spatial_mantel, comment="", message=FALSE, warning=FALSE}
correlog_co<-mantel.correlog(D.eco=dist(comm_co_clr_G121eO301A),D.geo=dist(metadata_co_G121eO301A[,c("Xcoord","Ycoord")]),nperm=999)
plot(correlog_co)
```

#### Distance decay plot
```{r colon_spatial_decay, comment="", message=FALSE, warning=FALSE}
toplot<-data.frame(spat_dist=as.numeric(dist(metadata_co_G121eO301A[,c("Xcoord","Ycoord")])),
                   comm_dist=as.numeric(dist(comm_co_clr_G121eO301A)))
ggplot(toplot,aes(x=spat_dist,y=comm_dist))+
  geom_point()+
  geom_smooth()
ggplot(toplot,aes(x=spat_dist,y=comm_dist))+
  geom_smooth()+
  xlim(c(0,8050))+
  xlab("Spatial distance (um)")+
  ylab("Aitchison distance")+
  theme_minimal()
anova(lmp(comm_dist~spat_dist,data=toplot))
summary(lmp(comm_dist~spat_dist,data=toplot))
```

#### Complex spatial structures

```{r colon_spatial_complex, comment="", message=FALSE, warning=FALSE}
set.seed(111)
G121eO301A_spat_SWNs<-listw.candidates(data.frame(metadata_co_G121eO301A[,c("Xcoord","Ycoord")]),
                                       nb = c("gab"), 
                                       weights = c("fdown", "fup"), 
                                       y_fdown = c(2,5), y_fup = c(0.1,0.5))
G121eO301A_W_sel <- listw.select(comm_co_clr_G121eO301A, G121eO301A_spat_SWNs, MEM.autocor = "positive", 
                                 p.adjust = T, method="FWD")
G121eO301A_W_sel$candidates
G121eO301A_W_sel$best$summary

par(mfrow=c(2,3))
for(i in 1:ncol(G121eO301A_W_sel$best$MEM.select)){
  s.value(metadata_co_G121eO301A[,c("Xcoord","Ycoord")],
          G121eO301A_W_sel$best$MEM.select[,i], 
          sub=G121eO301A_W_sel$best$summary$variables[i], 
          csub=2,
          include.origin = F,
          csize=0.8)
}
par(mfrow=c(1,1))
```

```{r colon_spatial_complex_plot, comment="", message=FALSE, warning=FALSE}
s.value(metadata_co_G121eO301A[,c("Xcoord","Ycoord")],
        rda(comm_co_clr_G121eO301A,
            G121eO301A_W_sel$best$MEM.select)$CCA$u[,1],
        sub=c("Spatial patterns of Colon"), 
        csub=2,
        include.origin = F,
        csize=0.8)

RsquareAdj(rda(comm_co_clr_G121eO301A,
               G121eO301A_W_sel$best$MEM.select))

```

## rlqESLTP of the Colon cryosection

```{r load_data_rlqESLTP, comment="", message=FALSE, warning=FALSE}
load("data/data.RData")
source("data/JEC_1743_sm_apps5.txt")
```

```{r prepare_data_rlqESLTP, comment="", message=FALSE, warning=FALSE}
comm_co_G121eO301A<-comm_co%>%
  as.data.frame()%>%
  rownames_to_column(var="microsample")%>%
  bind_cols(metadata_co)%>%
  filter(cryosection=="G121eO301A")%>%
  select(contains("bin_"))
metadata_co_G121eO301A<-metadata_co%>%
  filter(cryosection=="G121eO301A")

comp <- decostand(comm_co_G121eO301A,MARGIN=1,method="hellinger")	# Standardize composition by rows
colnames(comp)<-gsub("\\.",":",names(comp))
env <- data.frame(log_seq_counts=log(metadata_co_G121eO301A$fastp_total_after_filtering.total_bases),
                  div=rowSums(comp>0))			   			# Environmental matrix

genome_gifts<-genome_gifts[rownames(genome_gifts)%in%colnames(comp),]
genome_funct <- genome_gifts %>%
  as.data.frame() %>%
  rownames_to_column(var = "id") %>%                         # Preserve row names as a column
  pivot_longer(-id, names_to = "column", values_to = "value") %>%  # Reshape to long format
  mutate(group = substr(column, 1, 3)) %>%                  # Extract the first 3 characters of column names
  group_by(id, group) %>%                                   # Group by row (id) and prefix
  summarise(mean_value = mean(value, na.rm = TRUE), .groups = "drop") %>% # Calculate rowMeans for each group
  pivot_wider(names_from = group, values_from = mean_value) %>%  # Reshape back to wide format
  column_to_rownames(var = "id") 
genome_funct<-genome_funct[,-c(19:21)]

phy <- genome_tree				       		# Phylogenetic tree constructed with Phylomatic V3 + Phylocom

phy <- drop.tip(phy, setdiff(phy$tip.label, rownames(genome_funct)))
spa <- metadata_co_G121eO301A[,c("Xcoord","Ycoord")]				       		# Matrix of X Y spatial coordinates	
comp<-comp[,match(phy$tip.label,colnames(comp))]
genome_funct<-genome_funct[match(phy$tip.label,rownames(genome_funct)),]

mean(phy$tip.label==colnames(comp))

# Change the phylogenetic tree (an object of the class "phylo") into an object of the class "phylog", used by ade4
phylog <- newick2phylog(write.tree(phy)) # => "phylog" class
colnames(comp)<-gsub(":","_",colnames(comp))
rownames(genome_funct)<-gsub(":","_",rownames(genome_funct))
```

### Correspondence analysis of the composition matrix 

```{r correspondence_rlqESLTP, comment="", message=FALSE, warning=FALSE}
coacomp <- dudi.coa(comp, scan = FALSE, nf = 108)
summary(coacomp)
```

### Spatial analysis 

#### Gabriel graph

```{r gabriel_graph, comment="", message=FALSE, warning=FALSE}
# Create the Gabriel graph
nb1 <- graph2nb(gabrielneigh(as.matrix(spa)), sym = T)
nb1

# Spatial autocorrelation in diversity and sequence count distribution
sp.correlogram(nb1,log(env$div),order=8,method = "I")
sp.correlogram(nb1,env$log_seq_counts,order=8,method = "I")

lw1 <- nb2listw(nb1) # gives a neighbours list with spatial weights (i.e. matrix W)
lw1

```

```{r gabriel_graph_plot, comment="", message=FALSE, warning=FALSE}
par(mfrow=c(1,1))
plot(nb1, spa, pch = 21, bg = 'red')
title(main="Gabriel Graph")
class(nb1)

# [1] "nb"
# This is the Gabriel graph -the graph that in this case defines which points 
# are connected

# The matrix of spatial variables is obtained as the eigenvectors of a
# neighbour matrix. This matrix is analysed by principal component analysis.
nb1.neigh <- nb2neig(nb1)
vecspa <- scores.neig(nb1.neigh)
pcaspa <- dudi.pca(vecspa, coacomp$lw, scan = FALSE, nf = ncol(vecspa))
summary(pcaspa)

```

#### PCA analysis of the env matrix
```{r env_pca, comment="", message=FALSE, warning=FALSE}
pcaenv <- dudi.pca(env, row.w=coacomp$lw, scannf = FALSE, nf = 2)
summary(pcaenv)
```

#### The distances between species based on their biological traits, analyzed by PCoA.

```{r distances_traits, comment="", message=FALSE, warning=FALSE}
# Distance matrices for traits separately
listdis <- ldist.ktab(ktab.list.df(list(genome_funct)), c("Q"), scan = FALSE)
# choose 1

# Distance matrix for traits together
disT <- dist.ktab(ktab.list.df(list(genome_funct)), c("Q"), scan = FALSE)
# choose 1
pcotraits <- dudi.pco(disT, coacomp$cw, full = TRUE)
summary(pcotraits)
```

#### The distances between species based on their phylogenetic relatedness, analysed by PCoA.

```{r distances_phylogeny, comment="", message=FALSE, warning=FALSE}
pcophy <- dudi.pco(as.dist(as.matrix(phylog$Wdist)[names(comp),names(comp)]),coacomp$cw, full = TRUE)
summary(pcophy)

## 11. Tests for phylogenetic signals in traits
## *******************************************
#Function rtest.decdiv() in appendix S5, Pavoine et al., 2011. [1]
#All traits together
phystot <- rtest.decdiv(phylog, rep(1, 108), as.dist(as.matrix(disT)[names(phylog$leaves),names(phylog$leaves)]), 
                        nrep = 99, vranking = "droot", optiontest = "less",ties.method = "average", option = 3)
phystot

#All traits independently
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$D01)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$D02)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$D03)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$D05)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$D06)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$D07)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$D08)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$D09)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$B01)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$B02)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$B03)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$B06)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$B07)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$B08)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$B09)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)
rtest.decdiv(phylog, rep(1, ncol(comp)), as.dist(as.matrix(listdis$B10)[names(phylog$leaves),
                                                                        names(phylog$leaves)]), 
             nrep = 99, vranking = "droot", optiontest = "less", 
             ties.method = "average", option = 3)

## Update the PCOA of traits to remove traits without phylogenetic signal
# Distance matrix for traits together
disT <- dist.ktab(ktab.list.df(list(genome_funct[,-c(4,8,13)])), c("Q"), scan = FALSE)
# choose 1
pcotraits <- dudi.pco(disT, coacomp$cw, full = TRUE)
summary(pcotraits)
```

#### Extended RLQ analysis

```{r extended_rlq, comment="", message=FALSE, warning=FALSE}
rlqmix <- rlqESLTP(pcaenv, pcaspa, coacomp, pcotraits, pcophy, scan = F, nf = 2)
barplot(rlqmix$eig)
rlqmix$eig[1]/sum(rlqmix$eig)
# [1] 0.73 
rlqmix$eig[2]/sum(rlqmix$eig)
# [1] 0.12   
```

**1st axis**

```{r extended_rlq_axis1, comment="", message=FALSE, warning=FALSE}
plot(rlqmix, xy=spa, ax=1, wh="S")
plot(rlqmix, phy=phylog, ax=1, wh="P")
plot(rlqmix, traits=genome_funct[,-c(4,8,13)], ax=1, type="Q", wh="T")
plot(rlqmix, env=pcaenv$tab , ax=1, type="Q", wh="E")

s.value(metadata_co_G121eO301A[,c("Xcoord","Ycoord")],
        rlqmix$lR[,1],
        sub = "Patterns in S and E",
        csub = 1.5,
        include.origin = F,
        csize = 0.8)
dotchart.phylog(phylog, rlqmix$lQ[names(phylog$leaves), 1], cleav = 0, cdot = 1,
                scaling = F, yjoi = 0, cex.axis = 1.5, sub = "patterns in T and P", csub = 0)
```

```{r extended_rlq_axis1_overlay, comment="", message=FALSE, warning=FALSE}
library(jpeg)
library(grid)
library(cowplot)

spatial_image_coord <- read_csv("data/G121eO301A.csv") %>% 
  rename(microsample=ID)

# Load the JPEG image (ensure it's 1000x1000 pixels)
img <- readJPEG("data/G121eO301A.jpg")

img <- img[nrow(img):1, , ]  # This flips the image vertically

# Convert to rasterGrob
img_grob <- rasterGrob(img, width = unit(1, "npc"), height = unit(1, "npc"))

# Create the ggplot without the background
scatter_plot <- metadata_co_G121eO301A %>%
  left_join(spatial_image_coord, by = "microsample") %>%
  mutate(loading = rlqmix$lR[, 1]) %>%
  select(microsample, correlation, Xcoord_pixel_crop, Ycoord_pixel_crop) %>%
  ggplot(aes(x = Xcoord_pixel_crop+70, y = Ycoord_pixel_crop-20, fill = correlation)) +
  geom_point(shape=21, color="#333333") +
  scale_fill_gradient2(
    low = "#a10b47",      # negative
    mid = "white",    # zero
    high = "#edff47",   # positive
    midpoint = 0
  ) +
  coord_fixed(xlim = c(0, 1000), ylim = c(0, 1000), expand = FALSE) +
  theme_void()

# Combine image + plot
final_plot <- ggdraw() +
  draw_grob(img_grob, x = 0.05, y = 0, width = 0.85, height = 0.95) +  # Compress width to 90%
  draw_plot(scatter_plot, x = 0, y = 0, width = 1, height = 1)


# Display result
print(final_plot)
```

**2nd axis**

```{r extended_rlq_axis2, comment="", message=FALSE, warning=FALSE}
plot(rlqmix, xy=spa, ax=2, wh="S")
plot(rlqmix, phy=phylog, ax=2, wh="P")
plot(rlqmix, traits=genome_funct, ax=2, type="Q", wh="T")
plot(rlqmix, env=pcaenv$tab , ax=2, type="Q", wh="E")
```